---
title: "PredictivePenguins"
author: "Jeff Grayum"
date: "2/14/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries.
```{r}
library(tidyverse)
library(silgelib)
library(scales)
theme_set(theme_hc())
#install.packages("palmerpenguins")
library(palmerpenguins)
library(tidymodels)
```

Some quick exploratory analysis.
```{r}
penguins %>% 
  view() 

penguins %>%
  count(island, species)  

penguins %>%
  count(sex)

penguins %>%
  count(year)

penguins %>%
  filter(!is.na(sex)) %>%
  ggplot(aes(flipper_length_mm, bill_length_mm, color = sex, size = body_mass_g)) +
  geom_point(alpha = 0.7) +
  labs(x = "Flipper length (mm)",
       y = "Bill length (mm)",
       color = "Sex",
       size = "Body mass (grams)") +
  facet_wrap(~ species)
```

Let's build a model that can tell difference between male and female penguins.
```{r}
penguins_df <- penguins %>%
  filter(!is.na(sex)) %>%
  select(-year, -island)

set.seed(123)

#Split, train, test!
penguin_split <- initial_split(penguins_df, strata = sex)
penguin_train <- training(penguin_split)
penguin_test <- testing(penguin_split)

#We have a small dataset...  We'll create a resampled dataset via our training set to compute performance (bootstrapping of training data (resamples with replacement)!)  
set.seed(234)
penguin_boot <- bootstraps(penguin_train)
penguin_boot
```

Now lets make some models! We'll build two model specifications (glm and random forest)
```{r}
#We'll start with a basic logistic regression model.

glm_spec <- logistic_reg() %>%
  set_engine("glm")

#Let's also set up a random forest model. Set mode for classification and not regression.
rf_spec <- rand_forest() %>% 
  set_mode("classification") %>%
  set_engine("ranger")

```

We're going to use the dreaded workflow.  We need to add pre-processors to our models.
```{r}
penguin_wf <- workflow() %>%
  add_formula(sex ~ .)

glm_rs <- penguin_wf %>%
  add_model(glm_spec) %>%
  fit_resamples(
    resamples = penguin_boot,
    control = control_resamples(save_pred = TRUE, verbose = TRUE)
  )

rf_rs <- penguin_wf %>%
  add_model(rf_spec) %>%
  fit_resamples(
    resamples = penguin_boot,
    control = control_resamples(save_pred = TRUE, verbose = TRUE)
  )
```

Evaluate modeling
```{r}
collect_metrics(rf_rs)
collect_metrics(glm_rs) #slightly better!
```

#We saved our predictions.  Let's look at a confusion matrix!
```{r}
glm_rs %>%
  conf_mat_resampled()
```

Let's make some roc curves.
```{r}
#Dashed line is how our model would do randomly...further away our lines, the better.
glm_rs %>%
  collect_predictions() %>%
  group_by(id) %>%
  roc_curve(sex, .pred_female) %>%
  ggplot(aes(1 - specificity, sensitivity, color = id)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(show.legend = FALSE, alpha = 0.6, size = 1.2) +
  coord_equal() +
  labs(y = "Sensitivity")
```

Now it's time to come back to the testing data! Estimates performance on new data..
```{r}
penguin_final <- penguin_wf %>%
  add_model(glm_spec) %>%
  last_fit(penguin_split)

penguin_final

collect_metrics(penguin_final) #These are on the testing data (they're good!)

collect_predictions(penguin_final) %>%
  conf_mat(sex, .pred_class)
```

Lets extract the fitted workflow and tidy!
```{r}
penguin_final$.workflow[[1]] %>%
  tidy(exponentiate = TRUE) %>%
  arrange(estimate)
```

